---
title: "w6s1"
output: html_document
date: "2023-06-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(arrow)
```

## Open data file

I'm going to load the data I saved after the last exercise.
```{r load-data}

applications <- read_feather("~/Dropbox/McGill/teaching/2022-2023/ORGB690/data/app_data_starter.feather")

# lists all variables (columns) in the data
applications |> tbl_vars()
```
## What are gender shares at different levels?

First, let's pick TC 1600. 

```{r}

tc1600 <- applications |> 
  # keep only TC 1600, drop NA values for gender
  filter(tc==1600, !is.na(gender)) |> 
  group_by(examiner_id) |> 
  summarise(gender = first(gender, na_rm = TRUE), au = min(examiner_art_unit), .groups = "drop")

tc1600 |> count(gender) |> mutate(pct= n/sum(n))

```
### Overall share vs. perception of share

For women, we can average their perceptions of what share they represent in an AU and compare it to the overall share of women.

We'll have to create a share of women on the AU level, and then average that over all women.

```{r}

tc1600 <- tc1600 |> 
  group_by(au) |> 
  mutate(
    is_woman = if_else(gender == "female",1,0),
    share_women = sum(is_woman)/n()
    )

tc1600 |> ungroup() |> filter(gender=="female") |> summarise(avg_share = mean(share_women))
```

## Let's try for 1700
Same, as above, just for the different TC.


```{r}

tc1700 <- applications |> 
  # keep only TC 1600, drop NA values for gender
  filter(tc==1700, !is.na(gender)) |> 
  group_by(examiner_id) |> 
  summarise(gender = first(gender, na_rm = TRUE), au = min(examiner_art_unit), .groups = "drop")

tc1700 |> count(gender) |> mutate(pct= n/sum(n))

```
### Overall share vs. perception of share

For women, we can average their perceptions of what share they represent in an AU and compare it to the overall share of women.

We'll have to create a share of women on the AU level, and then average that over all women.

```{r}

tc1700 <- tc1700 |> 
  group_by(au) |> 
  mutate(
    is_woman = if_else(gender == "female",1,0),
    share_women = sum(is_woman)/n()
    )

tc1700 |> ungroup() |> filter(gender=="female") |> summarise(avg_share = mean(share_women))
```
